Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL

### 2020-12-11 main.tex
- paper changes

	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. Nevertheless, in the last years, various approaches have been studied to investigate their applicability in tuning operations. \\

@@

	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the hardware control system.\\

@@

	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon. For reason discussed later, the horizon varies in the applications between either 10 or 50 steps. An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}.\\

@@

	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
	As discussed in the previous sections, several tests were performed on the FERMI FEL.
	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility. 
	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments. Since there was a positive outcome, it is evident, that the utilised methods are at an operational level.\\
	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
	\subsection{MFRL tests}
		\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_training}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
		\caption{The verification of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_verification}
	\end{figure}

@@

@@

		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:AE-DYNA_verification}

@@

		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
		\caption{The training of the ME-TRPO on the FERMI FEL.}
		\label{fig:ME-TRPO_observables}
		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
		\caption{The verification of the ME-TRPO on the FERMI FEL.}
		\label{fig:ME-TRPO_verification}

@@

				\emph{AE-DYNA} &500 &  4.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
				\emph{ME-TRPO} &450&  5.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
				\emph{NAF} &1 074&  2.67 $\pm$  1.33 & -0.58 $\pm$  0.72 &0.00 $\pm$  0.02\\
				\emph{NAF2} &824& 2.56 $\pm$  1.26 & -0.66 $\pm$  0.80 &0.00 $\pm$  0.02\\

@@

 To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
 \vfill
 \newpage
 \vfill

>>>>

%	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. 
In the last years, various approaches have been studied to investigate their applicability in tuning operations. \\

@@

	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the control system.\\

@@

	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon. For reason discussed later, the horizon varies in the applications between either 10 or 50 steps. An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}. In our experiments the maximal intensity was provided by a manual optimization done of an expert. \\

@@

	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
	As discussed in the previous sections, several tests were performed on the FERMI FEL.
	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility. 
	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments. Since there was a positive outcome, it is evident, that the utilised methods are at an operational level.\\
	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
	\subsection{MFRL tests}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_training}
	\end{figure}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/Evolution_double.pdf}
		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_evolution_double}
	\end{figure}
	\begin{figure}
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Evolution_single.pdf}
	\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
	\label{fig:NAF_evolution_single}
\end{figure}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
		\caption{The verification of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_verification}
	\end{figure}

@@

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/Verification_DYNA_all_episodes.pdf}
		\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:AE-DYNA_verification}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
		\caption{The training of the ME-TRPO on the FERMI FEL.}
		\label{fig:ME-TRPO_observables}
	\end{figure}

@@

	\includegraphics*[width=0.5\textwidth]{Figures/Worst_episode_MBRL.pdf}
	\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
	\label{fig:Worst_episode_MBRL}

@@

%	\begin{figure}
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
%		\caption{The verification of the ME-TRPO on the FERMI FEL.}
%		\label{fig:ME-TRPO_verification}
%	\end{figure}

@@

				\emph{AE-DYNA} &500 &  3.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
				\emph{ME-TRPO} &450&  4.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
				\emph{NAF} &1 074&  2.56 $\pm$  1.96 & -0.66 $\pm$  1.22 &0.00 $\pm$  0.02\\
				\emph{NAF2} &824& 2.64 $\pm$  1.65 & -0.57 $\pm$  0.92 &0.00 $\pm$  0.03\\

@@

 To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.

 \section{Acknowledgements}
 Giulio Gaio
 \vfill
 \newpage
 \vfill
